What We Eval in the Shadows

### 2021-08-13 paper/main.tex
Clarify 5.3

Where does the expression passed to \eval come from? There are various means of
creating an expression, often these means correlate with a particular use case.
Our analysis records the values returned by some functions of interest as well
as their arguments. We build a direct acyclic graph that track the origin
of the values that are given to \eval. Fig~\ref{fig:provgraph} shows an example
of the origin graph for the expression \k{e}.
The origin of \k{e} includes both \k{parse} and \k{quote}. We pick an
origin by traversing the graph from the terminal leaf, \ie from the \eval node, following the left-hand side
member of assignments until reaching a root node. It represents the origin that is
further modified to yield the expression passed to \eval. In our example, the
origin is \k{parse}.

@@

\begin{itemize}[---]
\item {\it Reflection:} use of \k{match.call} to reflectively capture the
  expression that invoked this function, arguments are promises, and
  \k{substitute} is used to retrieve their source.
\item {\it String:} create from strings by invoking \k{parse},
  \k{str2lang}, or \k{str2expression}.
\item {\it Constructed:} invoked \k{quote} or \k{expression}, or by built with
  \k{call} or \k{as.call}.
\item {\it Environment: } created with \k{as.name} or \k{as.symbol}, typically
  to read a non-local environment.
\item {\it External: }  call to a C function. % Put an example
\end{itemize}

Table~\ref{tab:provenance} summarizes origins. Is shows frequencies by function
and by category. For technical reasons, we misclassify a small number of sites.
Manual inspection of numerous examples suggests that errors are rare. Some sites
are counted multiple times as they are invoked with different origins. Strings
correlate with dynamic code loading as done by \k{source} and \k{sys.source}.
Few calls (\packageNbParseFromFileSites in total) consume the result of calling
\k{parse} on a file. Most build strings programmatically.

\paragraph{Discussion}
The main use cases for \eval involve using \k{match.call}, \k{substitute}, or
\k{formals} to access and transform arguments of a function. Constructing
expressions from strings is also a common use case. The constructed category
shows that most \evals come from existing code that was modified by the
programmer before invoking \eval. Both constructed and reflection categories
roughly correspond to meta-programming. Some uses could be replaced by
macros if the R designers could be convinced to overcome their distaste for
those.

>>>>

Where does the expression passed to \eval come from? There are various means of
creating an expression, often these means correlate with a particular use case.
Our analysis records the values returned by some functions of interest as well
as their arguments. We build a direct acyclic graph that track the origin
of the values that are given to \eval. Fig~\ref{fig:provgraph} shows an example
of the origin graph for the expression \k{e}.
The origin of \k{e} includes both \k{parse} and \k{quote}, the starting nodes. We pick one origin by traversing the graph from the terminal leaf, \ie from the \eval node, following the left-hand side
member of assignments until reaching a starting node. It represents the origin that is
further modified to yield the expression passed to \eval. In our example, the
origin is \k{parse}.

@@

\begin{itemize}[---]
\item {\it Reflection:} use of \k{match.call} to reflectively capture the
  expression that invoked this function. Function arguments are promises, and
  \k{substitute} is used to retrieve their source. 
\item {\it String:} create from strings by invoking \k{parse},
  \k{str2lang}, or \k{str2expression}.
\item {\it Constructed:} invoked \k{quote} or \k{expression}, or by building with
  \k{call} or \k{as.call}.
\item {\it Environment: } created with \k{as.name} or \k{as.symbol}, typically
  to read a non-local environment.
\item {\it External: }  call to a C function. 
\end{itemize}

Table~\ref{tab:provenance} summarizes origins. It shows frequencies by function
and by category. For technical reasons, we misclassify a small number of sites.
Manual inspection of numerous examples suggests that errors are rare. Some sites
are counted multiple times as they are invoked with different origins. Strings
correlate with dynamic code loading as done by \k{source} and \k{sys.source}.
Few calls (\packageNbParseFromFileSites in total) consume the result of calling
\k{parse} on a file. Most build strings programmatically.

\paragraph{Discussion}
The main use cases for \eval involve using \k{match.call}, or \k{substitute} to access and transform arguments of a function. Constructing
expressions from strings is also a common use case. The constructed category
shows that most \evals come from existing code that was modified by the
programmer before invoking \eval. Both constructed and reflection categories
roughly correspond to meta-programming. Some uses could be replaced by
macros if the R designers could be convinced to overcome their distaste for
those.
